---
title: "COVis: covid-19 Visualizations"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
author: ""
output:
 rmarkdown::html_document:
 toc: true
 theme: cosmo
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,cache=F,autodep = T)

```
(Source code: https://github.com/b-lev/covid19-vis)

# Main plot

```{r init, echo=FALSE, include=FALSE}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(scales))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(ggrepel))
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(library(tidyquant))

theme_set(theme_gray(12))
theme_update(legend.position='top')
# read in JHU data
# transpose the colums/rows
# set type column to confirmed" (or deaths or recovered)
# rename columns


confirmed<- fread("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv") %>% 
	as.tbl %>%
	gather(date, val, -`Province/State`,-`Country/Region`,-Lat ,-Long) %>%
	mutate(date=mdy(date),type='confirmed') %>%
	rename('province'=`Province/State`,region=`Country/Region`) %>%
	replace_na(list(province=""))

deaths<- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv") %>% 
	gather(date, val, -`Province/State`,-`Country/Region`,-Lat ,-Long) %>%
	mutate(date=mdy(date),type='deaths') %>%
	rename('province'=`Province/State`,region=`Country/Region`)%>%
	replace_na(list(province=""))


recovered<- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_recovered_global.csv") %>% 
	gather(date, val, -`Province/State`,-`Country/Region`,-Lat ,-Long) %>%
	mutate(date=mdy(date),type='recovered') %>%
	rename('province'=`Province/State`,region=`Country/Region`)%>%
	replace_na(list(province=""))


tmp<- bind_rows(deaths,confirmed,recovered)%>% filter(val>0) %>%
	mutate(region=ifelse(region=='Korea, South',"S.Korea",region)) 

active<- inner_join(
	inner_join(filter(tmp,type=='confirmed'),
						 filter(tmp,type=='recovered'),
						 by=c('date','region','province','Lat','Long'),
						 suffix=c('.c','.r')),
	filter(tmp,type=='deaths'),
	by=c('date','region','province','Lat','Long')) %>%
	select(!contains('type')) %>%
	mutate(val=val.c-val.r-val,
				 type='active') %>%
	select(-c(val.c,val.r)) 

# bind all rows together
covid<- bind_rows(active,tmp)



# data scrapped manually from mass.gov. very different than JHU
ma.gov<-suppressMessages(read_csv('https://raw.githubusercontent.com/b-lev/massachusetts-covid19-report-archive/master/MA-stats.csv'))%>%
	mutate(date=mdy(date),
				 val=confirmed,
				 val.log=log(val),
				 rn=row_number()-2)%>% 
	mutate(delta=val-lag(val))

ct<- 
	suppressMessages(read_csv('https://covidtracking.com/api/states/daily.csv'))  %>%
	group_by(state) %>% arrange(state,date) %>%
	mutate(date=ymd(date)) %>%
	mutate(total=positive+negative) %>%
	select(-c(dateChecked))

start_val= 0.00004

```
Results below include up to `r max(ma.gov$date)` for MA 
[<a href="https://github.com/b-lev/massachusetts-covid19-report-archive">src</a>],
`r max(covid$date)` for countries 
[<a href="https://github.com/CSSEGISandData/COVID-19">src</a>], 
and `r max(ct$date)` for US states 
[<a href="https://covidtracking.com">src</a>].

Below is the most important plot. We want to be like S. Korea or lower.  This is a comparison of growth per capita, where  "day 0" is approx `r format(start_val*100,scientific=F,digits=4)`% of the country's population. This is a log scale, and so exponential/viral growth is a straight line.  The dashed grey lines show epidemic growth when the fraction of infected population doubles every 2 days, or 3 days, etc.  
```{r}


comparison<- filter(covid,type %in% c('confirmed'))  %>%
	group_by(region,date,type) %>%
	summarize(val=sum(val)) %>%
	group_by(region,type) %>%
	mutate(day=row_number())%>%
	filter( val>-10,region%in% c('Italy','US','S.Korea','Spain')) %>%
	mutate(pop=ifelse(region=='Italy',  60462000,
										ifelse(region=='US',  331003000,
													 ifelse(region=='Spain',47007367	,
													 			 ifelse(region=='S.Korea',  51269000)))))

comp.pc<- comparison %>%  ungroup %>% rowwise %>%
	mutate(valpc=val/pop, delta=abs(start_val-valpc)) %>%
	ungroup %>% 
	group_by(region) %>%
	filter(day>=which.min(delta)) %>%
	mutate(day=row_number()-1)

labels<- comp.pc %>% group_by(region) %>% filter(day==max(day))

template<- expand.grid(day=0:30,rate=c(2,3,4,5,10)) %>%
	as.tbl %>%
	mutate(region=paste(rate),valpc=start_val*2**(day/rate))

gr.labels=tibble(text=c('every\n2 days','every\n3 days','every\n4 days',
												'every\n5 days','every\n10 days'),
								 x=c(10,16,22,28,30),
								 y=c(.002,.002,.002,.002,.0003))

ggplot(comp.pc)+
	aes(x=day,y=valpc,group=c(region),color=region)+
	theme(legend.position = "none")+
	geom_line(size=1)+geom_point(size=1.5)+
	geom_line(data=template,linetype=2,color='grey')+
	geom_text(data=gr.labels,aes(label=text,group=NA,x=x,y=y),color='dark grey')+
	geom_text_repel(data=labels,aes(label=region),color='black',direction='x',seed =0)+
	scale_y_log10(breaks=round(filter(template,rate==2)$valpc,6),
								# label=scales::label_percent()
	)+
	scale_x_continuous(breaks=pretty_breaks(10))+
	coord_cartesian(ylim=c(start_val,.002))+
	labs(y='confirmed cases per capita (log scale)',
			 x=paste0('Days since ', 
			 				 format(start_val*100,scientific=F,digits=4),
			 				 '% of the population infected'))
```


# World data

This plot shows the spread of the virus over time across the world. We want the number of active cases  to drop to zero, the deaths to stay as low as possible, obviously, and the recovered cases to climb high. Confirmed is a cumulative count of everyone that has tested positive. 

```{r world, echo=FALSE,fig.width = 5,fig.height=3}
ggplot(covid %>%	 	group_by(date,type) %>% 
			 	summarize(val=sum(val)) %>%
			 	filter(type!='confirmed'))+
	aes(date,val,fill=reorder(type,desc(type)))+
	geom_bar(stat='identity')+
	scale_y_continuous(label=comma,breaks=pretty_breaks(),
										 sec.axis = sec_axis(~./7771316390,labels = percent))+
		labs(x=element_blank(),y=element_blank(),color=element_blank(),
				 fill=element_blank())+
	theme(legend.position=c(.2,.8))
```

Here's another view of the same data.

```{r, echo=FALSE,fig.width = 5,fig.height=3}


# All lines for the world
ggplot(covid %>% 
			 	group_by(date,type) %>% 
			 	summarize(val=sum(val)))+
	aes(date,val,group=type,color=type)+
	geom_line()+geom_point(size=.5)+
	scale_y_continuous(label=comma,breaks=pretty_breaks(),
										 sec.axis = sec_axis(~./7771316390,labels = percent))+
	theme(legend.position = c(.2,.7))+
	scale_x_date(breaks=pretty_breaks())+
	labs(x=element_blank(),y=element_blank(),color=element_blank())
```

These plots show how the numbers have grown over time in each country. All countries want to hit the bent curve that China has reported achieving. 

```{r world2}

tmp<- group_by(covid,region,type,date) %>% 
	filter(val>0,date>='2020-01-20') %>%
	summarize(val=sum(val)) %>%
	group_by(region) %>%
	filter(max(val)>=1)

labels<- group_by(tmp,region,type) %>% filter(date==max(date))  %>%
	ungroup

worst<- c(labels %>% 
						group_by(region) %>% 
						summarize(val=max(val)) %>% 
						top_n(6,val) %>% 
						pull(region) )

ggplot(filter(tmp,region %in% worst))+theme_gray(12)+
	aes(date,val,group=c(region),color=region)+
	scale_y_continuous(limits = c(0,NA),
										 label=comma,
										 breaks=pretty_breaks(8)	)+
	scale_x_date(breaks=pretty_breaks(10),
							 limits = c(min(tmp$date),max(tmp$date)+4))+
	geom_line()+geom_point(size=.5)+
	theme(legend.position="none")+
	facet_wrap(~type,scales = 'free',ncol=2)+
	theme(legend.position="none",
				axis.text.x = element_text(angle = 90, vjust = .5))+
	labs(x="",y='count')+
	geom_text_repel(data=labels %>%	filter(	(region %in% worst)| region=='US'),
									direction='y',
									segment.color='dark grey',
									hjust=0,
									seed=0,
									aes(label=region,x=date),color='black')+
		coord_x_date(xlim=c(min(tmp$date),max(tmp$date+18)))



```


# World Fatality Rates

Here is a plot of the fatility rate. It's really hard to tell whether this analysis makes sense given the unreliablity of the data overall. FWIW:

```{r fatality, echo=FALSE}
rate<-  	inner_join(confirmed %>% filter(date==max(date)) %>%  group_by(region) %>% summarize(confirmed=sum(val)),
										deaths %>%filter(date==max(date)) %>%  group_by(region) %>% summarize(deaths=sum(val)),
										by = c("region")) %>%
	# inner_join(recovered %>%filter(date==max(date)) %>%  group_by(region) %>%	summarize(recovered=sum(val)),
	# by = c("region")) %>%
	# filter(region!='Cruise Ship' ) %>%
	# mutate(region=ifelse(region=='Korea, South',"S.Korea",region)) %>%
	filter(deaths>0) %>% 
	mutate(rate=deaths/(deaths+confirmed)) %>%  ungroup %>% 
	rowwise %>%
	mutate(ci=1.96*sqrt(rate*(1-rate)/(deaths+confirmed))) %>%  ungroup %>% 
	group_by(region) %>% mutate(min_ci=min(ci)) 

ggplot(rate %>% filter(deaths>15))+
	aes(reorder(region,desc(rate)),rate)+
	scale_y_continuous(label=percent,breaks=pretty_breaks(10))+
	geom_bar(stat='identity')+
	geom_text(aes(label=deaths,y=.001),color='yellow',hjust=0,angle=90)+
	labs(x=element_blank(),y="fatality rate")+
	theme(legend.position="none",
				axis.text.x = element_text(angle = 90, vjust = .0,hjust=1))

```

# Massachusetts Data 

MA data from https://github.com/b-lev/massachusetts-covid19-report-archive. 

```{r MA.fit,fig.height=3,fig.width=5}

# Exponential fit
covid.fit<- function(actual,last=7,pred=7){
	# assumes columns: val.log, rn,date
	week<- top_n(actual,last,rn)
	start_date<- min(week$date)
	start_rn <- min(week$rn)
	lm<-with(week,lm(val.log~rn))
	fit.int<-lm$coefficients[['(Intercept)']]
	fit.s<- lm$coefficients[['rn']]
	fit<- tibble(rn=start_rn:(start_rn+last+pred)) %>% 
		mutate(date=start_date+rn-start_rn,
					 fit=fit.s*rn+fit.int,
					 val=exp(fit),
		) 
	return(fit)
}



# plot new cases per day
ggplot(ma.gov %>% select(date,val,delta) %>%na.omit)+
	aes(date,delta)+
	geom_bar(stat="identity")+
	scale_y_continuous(breaks=pretty_breaks(10))+
	labs(y="new cases in MA per day",x=element_blank())

# adjust "rn>0" to fit to different parts of the data
ma<-select(ma.gov ,date,val,val.log,rn) 
ma.fit<- covid.fit(ma,last=7,pred=5)

# # Show fit to line on log-scake
# ggplot()+
# 	geom_point(data=ma.gov ,aes(rn,val.log))+
# 	# geom_abline(aes(slope=fit.s,intercept = fit.int),linetype=2)+
# 	geom_line(data=ma.fit,aes(x=rn,y=fit),color='red')+
# 	scale_x_continuous(breaks=pretty_breaks())+
# 	scale_y_continuous(breaks=1:30)+
# 	# labs(title=paste(round(fit.s,3),"*x+",round(fit.int,3)))+
# 	labs(y="confirmed (log scale)",x="day")

ggplot()+
	geom_line(data=ma.fit,aes(x=date,y=val,color='prediction'),linetype=2)	+
	geom_point(data=ma.fit,aes(x=date,y=val,color='prediction'))	+
	geom_point(data=ma.gov %>% filter(date>='2020-03-01'),
						 aes(date,val,color='actual'))+
	geom_line(data=ma.gov %>% filter(date>='2020-03-01'),
						aes(date,val,color='actual'))+
	scale_y_continuous(breaks=pretty_breaks(5),label=comma,
										 sec.axis = dup_axis())+
	scale_x_date(breaks=pretty_breaks(5))+
	theme(legend.position = c(.2,.8))+
	labs(y="confirmed cases in MA",x="",color=element_blank())
```

Death rate in MA. 

```{r ma.death,fig.height=3,fig.width=5}
tmp<- ma.gov %>% filter(!is.na(Deaths)) %>%
	select(date,Deaths)  %>%
	mutate(val.log=log(Deaths),rn=row_number())

ggplot(ma.gov %>% filter(!is.na(Deaths)))+
	aes(date,Deaths,color='actual')+
	scale_x_date(breaks="3 days",date_labels = "%m/%d")+
	# scale_y_log10()+
	geom_point()+geom_line()+
	# geom_smooth(method='lm',aes(color='linear fit'))+
	geom_line(data=covid.fit(tmp,pred = 5),aes(y=val,color='exponential fit'),linetype=2)+
	geom_point(data=covid.fit(tmp,pred = 5),aes(color='exponential fit',y=val))+
	labs(color=element_blank())
```



The cumulative fraction of MA covid patients that are hospitalized.

```{r ma-hosp,fig.height=3,fig.width=5}
ma.hosp<-ma.gov %>% select(date,Hospitalized_yes,Hospitalized_no) %>%
	na.omit %>%
	mutate(hosp_frac=Hospitalized_yes/(Hospitalized_yes+Hospitalized_no))

ggplot(ma.hosp)+aes(date,hosp_frac)+
	geom_point()+geom_line()+
	scale_y_continuous(label=percent,limits=c(0,NA),
										 breaks=pretty_breaks(10))+
	labs(y="hospitalization in MA",
			 x=element_blank())


```

MA is performing an inadequate number of tests per day. 


```{r ma-testing, fig.height=3,fig.width=5}
ma.testing<-select(ma.gov,date,contains("Test_")) %>%
	gather(var,val,-date) %>% 
	na.omit %>%
	mutate(result=ifelse(str_ends(var,"_positive"),"pos","neg")) %>%
	group_by(result,date) %>%
	summarize(val=sum(val)) %>%
	filter(val>0) %>%
	mutate(delta=val-lag(val)) %>%
	na.omit

ggplot(ma.testing)+aes(date,delta,group=result,fill=result)+
	geom_bar(stat='identity',position = 'stack',width=1,color='black')+
	scale_y_continuous(breaks=pretty_breaks())+
	scale_x_date(breaks=pretty_breaks())+
	labs(y="Tests in MA per day")

```


```{r sheldon,fig.height=3,fig.width=5}
# tmp<- ma.testing %>%pivot_wider(names_from = result,values_from=c(val,delta)) 
# ggplot(tmp)+aes(date,delta_pos/(delta_pos+delta_neg))+
# 	geom_point()+geom_line()+
# 	scale_y_continuous(limits=c(0,NA),
# 										 label=percent,breaks = pretty_breaks())+
# 	scale_x_date(breaks=pretty_breaks())+
# 	labs(x="",y="Fraction of positive daily tests in MA")
```

```{r age,fig.height=3,fig.width=5}
ma.demo<- tibble(demo=c('0-19','20-29', '30-39', '40-49', '50-59', '60-69', '70-'),
								 pop=c(1549659,831628,809392, 972104, 919120, 627497, 599347))

ma.gov.deaths<- select(ma.gov,date,contains('age_')) %>%
	filter(date==max(date)) %>%
	pivot_longer(names_to = 'demo',names_prefix = "age_",
							 values_to = 'val',
							 values_drop_na = T,
							 cols = starts_with('age_'))
ggplot(inner_join(ma.demo,ma.gov.deaths,by='demo'))+
	aes(demo,val/pop)+
	geom_bar(stat='identity')+
	scale_y_continuous(label=percent,breaks = pretty_breaks())+
	# theme(axis.text.x = element_text(angle=0,hjust=1))+
	geom_text(aes(label=val,y=val/pop*.9),color='yellow')+
	labs(x='age demographic',y='Fraction of that demo\'s population',title='MA Confirmed Cases')
```

```{r death-demo,fig.height=3,fig.width=5}


ma.deaths<-suppressMessages(
	read_csv('https://raw.githubusercontent.com/b-lev/massachusetts-covid19-report-archive/master/deaths.csv')
)%>%
	mutate(date=mdy(date))

ggplot(ma.deaths)+
	aes(reorder(age,as.integer(str_sub(age,end=-2))),group=gender,fill=gender)+
	geom_bar(position='stack')+
	scale_y_continuous(breaks = pretty_breaks(10))+
	theme(legend.position = c(.2,.7))+
	labs(title='MA Covid deaths',x='age group',y='count',fill=element_blank())
	
	
	
```

```{r counties,fig.height=6,fig.width=5}

ma.county.pop<-read_csv("region,pop\nMiddlesex,	1614714\nWorcester ,	830839\nSuffolk ,	807252\nEssex ,	790638\nNorfolk ,	705388\nBristol ,	564022	\nPlymouth 	,518132	\nHampden ,	470406\nBarnstable ,	213413\nHampshire ,	161355	\nBerkshire ,	126348	\nFranklin ,	70963	\nDukes and Nantucket 	,28679")

ma.counties<- select(ma.gov, c(date,Barnstable,	Berkshire,	Bristol,'Dukes and Nantucket',	Essex,	Franklin,	Hampden	,Hampshire, Middlesex, Norfolk, Plymouth,Suffolk, Worcester))  %>% 
	gather(region,val,-date) %>% na.omit() %>%
	inner_join(ma.county.pop,by='region') %>%
	filter(val>0) %>%
	mutate(valpc=val/pop,day=row_number()) 

ggplot(ma.counties)+
	aes(date,val,group=region,color=reorder(region,desc(val)))+
	scale_y_log10(label=comma)+
	scale_x_date(breaks=pretty_breaks(8))+
	geom_point()+geom_line()+
	annotation_logticks() +
	labs(y='Count of Confirmed Cases (cumulative)',
			 color=element_blank(),
			 x="")

start_val=4e-05

ma.counties.f<- ma.counties%>% ungroup %>% rowwise %>%
	mutate( delta=abs(start_val-valpc)) %>%
	ungroup %>% 
	group_by(region) %>% 
	filter(day>=which.min(delta)-1,valpc>start_val) %>%
	mutate(day=row_number()-1,max_valpc=max(day))


template<- expand.grid(day=0:18,rate=c(2,3,4,5,10)) %>%
	as.tbl %>%
	mutate(region=paste(rate),valpc=start_val*2**(day/rate))

gr.labels<- template %>%group_by(region) %>%
	filter(day==max(day)-1) %>% mutate(text=paste('every',rate,'\ndays'))


ggplot(ma.counties.f)+
	aes(x=day,y=valpc,group=c(region),color=reorder(region,max_valpc))+
	theme(legend.position = "none")+
	geom_line(size=1)+geom_point(size=1.1)+
	geom_line(data=template,linetype=2,color='grey')+
	geom_text(data=gr.labels,aes(label=text,group=NA),color='dark grey')+
	# geom_text_repel(data=labels,aes(label=var),color='black',direction='x',seed =0)+
	scale_y_log10(breaks=filter(template,rate==2)$valpc,label=percent)+
	scale_x_continuous(breaks=pretty_breaks(10))+
	# coord_cartesian(ylim=c(start_val,.01))+
	labs(y='confirmed cases per capita in county',
			 color="",
			 x=paste0('Days since ', 
			 				 format(start_val*100,scientific=F,digits=4),
			 				 '% of the population infected'))+
	theme(legend.position = 'top')


```

# US States


Deaths in the US.

```{r us.deaths,fig.height=3,fig.width=5}
tmp<- covid %>% filter(region=='US',type=='deaths',val>100) %>%
	select(date,val)  %>%
	mutate(val.log=log(val),rn=row_number())
# 
# tmp<- inner_join(	deaths %>% filter(region=='US',date>='2020-03-21'),
# 	confirmed %>% filter(region=='US',date>='2010-03-21'),
# 	by=c("province", "region", "Lat", "Long", "date"),
# 	suffix=c('.d','.c') ) %>% 
# 	mutate(val=val.d/(val.d+val.c)) %>%
# 	mutate(val.log=log(val),rn=row_number())
# 	

ggplot(tmp )+
	aes(date,val,color='actual')+
	scale_x_date(breaks="3 days",date_labels = "%m/%d")+
	scale_y_continuous(breaks=pretty_breaks(10))+
	geom_point(size=2)+geom_line(size=1)+
	# geom_smooth(method='lm',aes(color='linear fit'))+
	geom_line(data=covid.fit(tmp,pred = 5),aes(y=val,color='exponential fit'),linetype=2)+
	labs(y="Deaths in US",x="",color="")+
	geom_point(data=covid.fit(tmp,pred = 5),aes(color='exponential fit',y=val))

```

(Based on https://covidtracking.com/api/states/daily.csv.)

````{r states,fig.height=9,fig.width=7}


state.pops<-suppressMessages(read_csv('state-pop-census.csv')) %>%
	mutate(state=state.abb[match(state_name,state.name)]) %>%
	mutate(state=ifelse(state_name=="DC","DC",state)) %>%
	select(state,pop)

testing<-ct %>% group_by(state) %>% arrange(state,date) %>%
	mutate(delta=total-lag(total,1)) %>% 
	inner_join(state.pops,by='state')	%>% ungroup %>%
	mutate(total_per_cap=total/pop) %>% 
	rename(confirmed=positive)


tmp<-select(testing,date,state,confirmed,death,hospitalized) %>% gather(type,val,-date,-state)%>%
	filter(date>'2020-02-23')
labels<- group_by(tmp,type) %>% 
	filter(date==max(date)) %>%
	top_n(5,val)

ggplot(tmp %>% filter(state %in% labels$state) %>% filter(!is.na(val)) )+
	aes(date,val,group=state,color=state)+
	scale_y_continuous(breaks=pretty_breaks(8), limits = c(0,NA),	)+
	scale_x_date(breaks=pretty_breaks(10),
							 date_labels = "%b%d")+
	geom_line()+geom_point()+
	theme(legend.position="none",
				axis.text.x = element_text(angle = 90, vjust = .5),
				panel.spacing = unit(2, "lines"))+
	geom_label_repel(data=labels,direction='x',label.padding = .05,label.size = 0,
						aes(label=state,x=date+5),
						color='black',
						position = position_jitter(width = 5,height = .1,seed=1)
	)+
	facet_wrap(type~.,nrow=2,scales = 'free')+
	labs(x="",y='')


```

# US Testing Statistics

Testing in each state.  Sometimes, health care workers are tested, to be sure they are not infected. That is one explanation for why some states show many negatives.  (Based on https://covidtracking.com/api/states/daily.csv.)


```{r testing, fig.height=7}


tmp<- filter(testing,date==max(date))  %>%
	mutate(rate=confirmed/(confirmed+negative))
tmp<- mutate(tmp,	 st.f=factor(state,levels =tmp$state[order(tmp$rate)]))

ggplot(tmp)+
	aes(y=st.f,x=rate)+
	geom_bar(stat='identity')+
	scale_x_continuous(label=percent,breaks=pretty_breaks())+
	geom_text(aes(label=prettyNum(confirmed+negative,big.mark=','),x=round(rate+.15,1)),hjust=1)+
	labs(x='COVID testing positive rate  (total tests)',y="State")



```

This plots show how long it is taking each state to test its population, on a per capita basis. 
(I'm assuming, optimistically, that the test had not be given to the same person twice.)

```{r testing2}

## tests per capita
labels=filter(testing,date==max(date)) %>% ungroup %>%
	top_n(n=25,wt=total_per_cap)

ggplot(testing %>% select(date,total_per_cap,state) %>% na.omit)+
	aes(date,total_per_cap,group=state,color=state)+
	geom_line()+
	theme(legend.position = 'none',
				axis.text.x = element_text(angle = 90, vjust = .5))+
	geom_text_repel(data=labels,aes(label=state,x=date+8),
									nudge_x=1,
									seed=0,
									direction='x')+
	scale_x_date(breaks=pretty_breaks(10))+
	scale_y_continuous(label=percent,breaks=pretty_breaks(10))+
	labs(y="Cumulative perc. of state population tested",x="")


```

```{r weekday-drop}
# 
# tmp<- ct %>% filter(!is.na(death),death>0) %>% group_by(date,day=paste(wday(date),weekdays(date))) %>% summarize(val=sum(death)) %>%
# 	ungroup %>%
# 	mutate(delta=val-lag(val,1),ratio=delta/lag(delta,1))
# 
# ggplot(	tmp)+
# 	aes(day,ratio)+
# 	scale_x_discrete()+
# 	scale_y_continuous(limits = c(0,NA))+
# 	# geom_boxplot(notch=F)+
# 	geom_point(size=2)

```


Thanks to JHU for the data (https://github.com/CSSEGISandData/COVID-19). Results are only as good as the data, which is only as good as the amount of testing done. 

I hereby disclaim any and all representations and warranties with respect to this data, including accuracy, fitness for use, and merchantability. Reliance for medical guidance or use in commerce is strictly prohibited.
